---
title: "7st Blog Post - Group 9"
author: "Group 9"
date: "2024-12-07"
date-modified: "2024-12-07"
draft: FALSE
---

First, we decided to continue exploring other factors that can affect income,
seeing how income has changed over years. Since we have not seen and analyzed 
this important factors. 

The tentative thesis we decided to choose is that the condition for the minority
group and the groups needing more care from are not great through the years. 
This could be discussed for several groups, including veteran groups and 
minority races, and from several aspects, including average income, 
employment rate, and the current revenue of firms with more groups needing help,
and the change of amount of new opening business by that group. Under this 
paragraph is a plot of change of average income through year. From here,
the condition doesn't seem so well for minority groups. Although the average
income for non-minority is having an increasing trend, it doesn't increase as 
much for minority and there sure needs something to be done. We have also made a 
plot for how the employees of each race have changed through year in blogpost 4,
which doesn't seem so well for minority groups as well.


```{r, echo = FALSE, results = 'hide', message = FALSE}
library(tidyverse)
data <- readRDS(here::here("dataset/cleaned_data_W_revenue_combined.rds"))
attach(data)
```
```{r}
data |>
  mutate(average_income = PAYANN / EMP * 1000) |>
  filter(VET_GROUP_LABEL == "Total",
         NAICS2017_LABEL == "Total for all sectors",
         ETH_GROUP_LABEL == "Total",
         SEX_LABEL == "Total") |>
  subset(select = c(average_income, YEAR, RACE_GROUP_LABEL, YIBSZFI_LABEL)) |>
  group_by(RACE_GROUP_LABEL) |>
  filter(RACE_GROUP_LABEL %in% c("Nonminority", "Minority", "Equally minority/nonminority")) |>
  filter(YIBSZFI_LABEL %in% c("Firms with less than 2 years in business", "Firms with 16 or more years in business","Firms with 4 to 5 years in business")) |>
  ggplot() +
  geom_line(aes(x = YEAR, y = average_income, color = RACE_GROUP_LABEL, linetype = YIBSZFI_LABEL)) +
  labs(color = 'Race category', 
       linetype = "Years in business") +
  ylab("Average income / $")
```


## Plan to improve Visualizations:
### Titles and Captions:
Use descriptive titles that summarize the key insight of each graph (e.g., "Trends in Average Employment by Racial Groups (2017–2021)").

Add captions to provide context for each visualization, explaining the data source, key variables, and the main takeaway.

### Annotations:
Highlight key data points or trends with annotations. For example, mark significant peaks or declines in a line graph and explain them briefly (e.g., "2017 Peak due to policy changes").

### Highlights:
Use color schemes to emphasize important trends or categories. For instance, consistent colors for specific racial groups across all graphs will help in comparison.

Employ ggpubr functions like ggpar() to refine axis labels, text sizes, and legends for clarity.

### Polished Designs:
Utilize ggpubr for adding statistical comparisons (e.g., stat_compare_means()) to support data interpretation.
Add gridlines or adjust themes for better visual appeal using theme_pubr().

## Tables:

### Title and Context:
Include clear, bold titles for each table summarizing its content (e.g., "Annual Payroll Trends Across Demographics (2017–2021)").

Add concise footnotes explaining data sources or column definitions.

### Formatting with GT:
Use the gt package to format tables professionally. Features like grouping rows, adding summary rows, and highlighting columns can improve readability.

Add highlights for significant figures, such as changes in revenue or employment disparities.

## Next Steps:
Apply a consistent style across all visualizations and tables, focusing on simplicity and accessibility.

Review for alignment between visualizations and narrative, ensuring that every figure and table directly supports your analysis and conclusions.

Validate design choices with peers or advisors for clarity and impact.
