---
title: "7st Blog Post - Group 9"
author: "Group 9"
date: "2024-12-07"
date-modified: "2024-12-07"
draft: FALSE
---

First, we decided to continue exploring other factors that can affect income,
seeing how income has changed over years. Since we have not seen and analyzed 
this important factors. 

The tentative thesis we decided to choose is that the condition for the minority
group and the groups needing more care from are not great through the years. 
This could be discussed for several groups, including veteran groups and 
minority races, and from several aspects, including average income, 
employment rate, and the current revenue of firms with more groups needing help,
and the change of amount of new opening business by that group. Under this 
paragraph is a plot of change of average income through year. From here,
the condition doesn't seem so well for minority groups. Although the average
income for non-minority is having an increasing trend, it doesn't increase as 
much for minority and there sure needs something to be done. We have also made a 
plot for how the employees of each race have changed through year in blogpost 4,
which doesn't seem so well for minority groups as well.


```{r, echo = FALSE, results = 'hide', message = FALSE}
library(tidyverse)
data <- readRDS(here::here("dataset/cleaned_data_W_revenue_combined.rds"))
attach(data)
```
```{r}
data |>
  mutate(average_income = PAYANN / EMP * 1000) |>
  filter(VET_GROUP_LABEL == "Total",
         NAICS2017_LABEL == "Total for all sectors",
         ETH_GROUP_LABEL == "Total",
         SEX_LABEL == "Total") |>
  subset(select = c(average_income, YEAR, RACE_GROUP_LABEL, YIBSZFI_LABEL)) |>
  group_by(RACE_GROUP_LABEL) |>
  filter(RACE_GROUP_LABEL %in% c("Nonminority", "Minority", "Equally minority/nonminority")) |>
  filter(YIBSZFI_LABEL %in% c("Firms with less than 2 years in business", "Firms with 16 or more years in business","Firms with 4 to 5 years in business")) |>
  ggplot() +
  geom_line(aes(x = YEAR, y = average_income, color = RACE_GROUP_LABEL, linetype = YIBSZFI_LABEL)) +
  labs(color = 'Race category', 
       linetype = "Years in business") +
  ylab("Average income / $")
```