---
title: "4th blog post group 9"
author: "Group 9"
date: "2024-11-4"
date-modified: "2024-11-4"
draft: FALSE
---

While cleaning the data, there was one specific variable, the revenue, which has way more NA values in it than any other variables. However, as a very significant 
numerical variable, it would be a pretty big loss to the data set if we really completely ignore this variable and let it go. As a result, two sets of cleaned 
data was presented, one with the revenues presented but as a result of data cleaning, there are only 16879 data points left, and one without without the revenue 
variable to provide more information about all other variables, and as a result, has 43575 data points left.

After all works done, the data has seven categorical variables, which includes the sector of industry of the company, the sex label, the ethical group label, the 
race label, the veteran label, the time company has been in the industry and the year that data was collected. There are also three numerical variables including 
the number of employers, the number of employees and their payrolls. The standard error for those variables are also included. In one of the data sets, the 
revenue and standard error for the revenue is also included.

```{r}
#| output: FALSE
data <- readRDS(here::here("dataset/cleaned_data_N_revenue.rds"))

data
# Basic data exploration
str(data)
summary(data)

## Average Employment Trends by Race

library(ggplot2)
library(dplyr)
library(scales)

# Function to format numbers in millions
format_m <- function(x) {
  paste0(round(x/1000000, 1), "M")
}
```
```{r}
# Employment trends by race
data %>%
  filter(!RACE_GROUP_LABEL %in% c("Total", "Unclassifiable", "Classifiable")) %>%
  group_by(YEAR, RACE_GROUP_LABEL) %>%
  summarise(avg_employment = mean(EMP), .groups = 'drop') %>%
  ggplot(aes(x = YEAR, y = avg_employment, color = RACE_GROUP_LABEL)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3, alpha = 0.8) +
  theme_minimal() +
  labs(title = "Average Employment Trends by Race",
       subtitle = "Excluding Total, Unclassifiable, and Classifiable categories",
       x = "Year",
       y = "Average Number of Employees (M)",
       color = "Race") +
  scale_y_continuous(labels = format_m) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 8),
        panel.grid.minor = element_blank())
```
In this table we are able to see the average employment trends by race. The trend starts in 2017 and ends in 2021. While most race categories during this time 
remain below the 1 million average number mark, the White and Non-minority race categories quickly rise in the first year and remain fairly stagnant at about the 
4 million average employee mark. All other races either remain very low in average employee count or grow slightly. These other numbers also stagnated at their 
lower average numbers. 

```{r}

library(tidyverse)

# Load your data
data <- read_rds("C:/Users/loveh/Downloads/cleaned_data_N_revenue (1).rds") 
data_long <- data %>%
  group_by(SEX_LABEL) |>
  select(SEX_LABEL, YEAR, PAYANN)

# Select relevant columns for plotting
ggplot(data_long, aes(x = YEAR, y = PAYANN))+
  geom_col() +
  labs(
    title = "Annual Pay Trends by Sex",
    x = "Year",
    y = "Annual Pay (PAYANN)",
    color = "Sex"
  ) +
  theme_minimal()

```

The bar chart shows the trend in annual pay from 2017 to 2021, with a notable peak in 2017, followed by a sharp decrease in 2018, and relative stability from 2018 
to 2021. The initial high level in 2017 suggests that this year had an unusual spike in earnings, significantly above the levels seen in subsequent years.

From 2017 to 2018, there was a significant drop in annual pay. This decline might be due to one-time factors that impacted 2017, such as special bonuses, 
incentives, or policy changes that increased earnings temporarily. Additionally, shifts in economic conditions or organizational budget adjustments could explain 
the substantial decrease seen in 2018.

Between 2018 and 2021, the annual pay stabilized, with only minor fluctuations. This steady trend may reflect a period of economic balance or consistent salary 
structures, suggesting that the organization or industry aimed for controlled spending and sustainable pay levels. Economic factors, such as inflation adjustments 
or policy regulations, might have contributed to maintaining this stable pay pattern.

Overall, the data indicates that after the spike in 2017, annual pay was adjusted to a more sustainable level in subsequent years. This pattern suggests possible 
restructuring or cost-cutting measures aimed at achieving long-term financial stability.
